---
title: "STAT401_HW3"
author: "김정현"
output: word_document
---
# Q1.
```{r}
options(digits=4)
mu <- c(2, -1)
sigma <- matrix(c(2, 2, 2, 5), byrow = TRUE, nrow = 2)
eigen_result <- eigen(sigma)
```

## (a)
```{r}
a11 = round(eigen_result$vectors[1, 1],4)
a12 = round(eigen_result$vectors[2, 1],4)
a21 = round(eigen_result$vectors[1, 2],4)
a22 = round(eigen_result$vectors[2, 2],4)

cat(paste("PC 1:",a11, "x_1 + ", a12, "x_2"))
cat("\n")
cat(paste("PC 2:", a21, "x_1 + ", a22, "x_2"))
```

First PC: $Y_1 = 0.4472X_1 + 0.8944X_2$
Second PC: $Y_1 = -0.8944X_1 + 0.4472X_2$

## (b)
```{r}
round(eigen_result$values[1]/sum(eigen_result$values),4)
```
The proportion of total population variance explained by first PC is 0.8571.


## (c)
```{r}
i = 2
j = 1
cat("Correlation between X_2 and Y_1 : ",sqrt(eigen_result$values[j])*a12 / sqrt(sigma[i,i]))
```

Correlation between $X_2$ and $Y_1$ is 0.9798.

## (d)
```{r}
obs <- matrix(c(3,1), byrow = T, nrow = 2)
a1 <- matrix(c(a11, a12), byrow = F, nrow = 1)
a1 %*% (obs-mu)

```
The PC score for first principal component is 2.236 (same with $\sqrt{5}$).

## (e)
```{r}
pc1 <- mu[2] - a12/ a11 * mu[1]
pc2 <- mu[2] - a22/ a21 * mu[1]
plot(mu[1],mu[2], xlim=c(-4,4), ylim=c(-3,3), xlab='X1', ylab='X2')
abline(h=-1); abline(v=2)
abline(pc1, a12 / a11, col='blue')
abline(pc2, a22 / a21, col='red')
text(mu[1]+0.3, 3, "X1") ;text(4, mu[2]+0.3, "X2")
text(4, 2, "Y1", col='blue'); text(-3, 1, "Y2", col='red')
```
Since $\lambda_1$ is the largest eigenvalue, the major axis lie on direction $a_1$ and the minor axis lie on direction $a_2$.



# Q2.
```{r}
# Mean vector
x_bar <- c(95.5, 164.4, 55.7, 93.4, 18.0, 31.1)

# Covariance matrix
S <- matrix(c(3266, 1344, 732, 1176, 163, 238,
              1344, 722, 324, 537, 80, 118,
              732, 324, 179, 281, 39, 57,
              1176, 537, 281, 475, 64, 95,
              163, 80, 39, 64, 10, 14,
              238, 118, 57, 95, 14, 21), nrow = 6, byrow = TRUE)
```

## (a)
```{r}
q2.pca.cov = princomp(covmat = S)
summary(q2.pca.cov)
```
The data can be summarized by 1 dimension (with cumulative proportion approximate 95.8%),  which is smaller than 6 dimensions.

## (b)
```{r}
std_devs <- sqrt(diag(S))

R <- matrix(nrow = 6, ncol = 6)

for (i in 1:6) {
  for (j in 1:6) {
    R[i, j] <- S[i, j] / (std_devs[i] * std_devs[j])
  }
}

rownames(R) <- colnames(R) <- c("X1", "X2", "X3", "X4", "X5", "X6")

print(R)
summary(princomp(covmat = R))
```
The data can be summarized by 1 dimension (with cumulative proportion approximate 94.3%), which is smaller than 6 dimensions.

## (c)
The proportion of total variance has similar value.
However, the eigen values and eigen vector are different since scaling is made in correlation matrix.


# Q3.
## (a)
```{r}
radio <- read.table('radiotherapy.dat', header = TRUE, sep = "")[,-6]
sapply(radio, var)
```

Since the variance has difference in scale (especially in Symptoms), it is better to use correlation matrix R.

## (b)
```{r}
q3.pca = prcomp(radio, scale = T)
round(q3.pca$rotation, 3)
```

## (c)
```{r}
summary(q3.pca)
```
1) By total proportion of variance: Since adding second proportion explains over 70% variance, choosing 2 principal components is appropriate.

2) By using scree plot: By drawing the scree plot, choosing 2 principal components is appropriate.

```{r}
screeplot(q3.pca, type = "l")
```

3) By choosing variance larger than 1: Choosing 1 principal component is appropriate.

Therefore, using 2 principal component is appropriate.

## (d)
1) First principal component explains the overall effect of 5 variables.

2) Second principal component: By considering the absolute value of coefficients larger than 0.2, exclude Food-consumed and appetite. Then, the second principal component can be interpreted as active reason (Symptom, Activity) against non-active reason (Sleep).

## (e)
Since 2 PC explain 73.95% of total sample variance, the data is summarized with 2 PC given in this data.